Tackling discrepancies in freshwater invertebrate trait databases and taxonomic resolution: Harmonising across continents and trait aggregation

### 2021-01-26 Paper/Draft_harmonization_paper.tex
26.01: Fixed references error
26.01.: Final fixes before send to Brooke Pt.1

% From freshwater biology:
% List all sources in the reference list alphabetically by name. In text citations should follow the author-date method. This means that the author's last name and the year of publication for the source should appear in the text, for example, (Jones, 1998), and a complete reference should appear in the reference list at the end of the paper.

@@

% In press or forthcoming: van Bergen, E., de Jong, P. F., Maassen, B., Krikhaar, E., Plakas, A., & van der Leij, A. (in press). IQ of four-year-olds who go on to develop dyslexia. Journal of Learning Disabilities. DOI: 10.1177/0022219413479673

@@

2. We describe discrepancies in the definitions of traits used to create freshwater invertebrate trait databases in Europe, North America, New Zealand, and Australia. Based on our comparisons of these trait databases, we established 4 novel trait datasets by harmonising trait definitions of commonly used traits. Two of these datasets were used to compare aggregated traits obtained by different aggregation methods with traits assigned by experts at family-level. We compared trait aggregation that were based on the mean and the median respectively, each with different weightings. Furthermore, we explored the effect of harmonisation and trait aggregation by re-analysing data from a case study.

@@

We extracted information from 6 trait databases from Europe, North America, Australia, and New Zealand and harmonised 7 grouping features. Trait information for Europe was obtained from the freshwaterecology.info database (\cite{schmidt-kloiber_www.freshwaterecology.info_2015}) and the Tachet database (\cite{ussegliopolatera_biological_2000}). The freshwaterecology.info contains taxa on species-level, while taxa recorded in the Tachet database are on species, genus and family-level. Trait information for North America was obtained from \citet{twardochleb_trait_data_2020} and complemented by \citet{vieira_database_nodate}. Data on body form for European and North American taxa was based on expert knowledge (\cite{polatera_personal_information_2020}). For Australia and New Zealand, we used trait databases from \citet{kefford_integrated_2020} and \citet{Philips_and_Smith_NZ_DB_2018} respectively. To increase readability we refer to the databases as well as the datasets we extracted from them by the name of the continent they originate from, except for European databases, which we refer to by their commonly used names (freshwaterecology.info and Tachet database).

@@

    \textit{a} The trait "bluff (blocky)" occurred in the North America database (Vieira et al.) and was newly classified by expert knowledge into cylindrical and flattened (\cite{polatera_personal_information_2020}). \\  
    \textit{b} Reflects the different size classifications by the North American trait databases and the other trait databases. \\

@@

To examine the influence of the taxonomic hierarchy and the trait variability on the outcome of the different trait aggregation methods we created three hypothetical examples of different taxonomic hierarchies (scenarios). 
1) A family with an equal number of genera and species (5 genera each with 5 species respectively), denoted as \textit{sim\_base}.
2) A family where one genus has a much larger number of species than the other 4 genera (1 genus with 13 species, 4 genera with 3 species respectively), denoted as \textit{sim\_extreme}. 
3) A family where all genera have a different number of species (8, 2, 7, 3, 5), denoted as \textit{sim\_variation}. In total every family consisted of 25 species. To each scenario a hypothetical grouping feature with 3 traits was assigned. The 25 affinities for each trait were simulated by sampling from a truncated normal distribution with a mean value of 0.5 and 5 levels of standard deviations (0.2, 0.4, 0.6, 0.8, and 1) respectively to simulate different levels of trait variability. The truncated normal distribution was bound to 0 and 1. Simulated trait affinities were converted to percentages, similar to the data processing of the trait databases. The sampling was repeated 100 times for each standard deviation. Hence, in total 100 datasets for each level of trait variability were simulated, resulting in 12500 simulated trait affinities for each simulated trait ($25$ species per family $* 5$ levels of trait variability $* 100$ replicates). These 12500 simulated affinities per trait were assigned to each scenario. The above mentioned trait aggregation methods were applied to each simulated dataset. 
% ...resulting in 2500 aggregated affinities per trait and scenario ($12500$ trait affinities per trait divided by $25$, the number of species and multiplied by $5$, the number of aggregation methods). 
The results were compared based on the produced range of aggregated trait affinities between levels of trait variability and scenario. We also compared for each simulated dataset the differences in trait affinities obtained by each aggregation method.

@@

\subsection*{Comparison of family-level aggregated traits with family-level assigned traits}

@@

For the North American dataset, we compared aggregated trait affinities with assigned trait affinities on family-level for the grouping features feeding mode, respiration, size, voltinism, and locomotion. The assigned trait affinities at family-level are part of the North American database (Twardochleb et al.) and originate from expert knowledge. Trait information was available for 94 families of which all were present in the North American dataset (total number of cases 1598). The traits were on the categorical scale and were converted to binary traits prior to the comparison with aggregated trait affinities.

@@

Definitions of grouping features and traits varied in their level of detail in the investigated trait databases. The freshwaterecology.info database, the Tachet database and the North American database (Twardochleb et al.) provided more detailed descriptions of their trait information compared to the North American (Vieira et al.) and New Zealand database. The Australian trait database (\cite{kefford_integrated_2020}) is a collection of seven specific trait datasets. Thus, grouping features occur multiple times with varying differentiation into traits. Depending on the dataset trait information is described with more or less detail.

The definition of grouping features varied across databases mainly concerning their differentiation into traits but also in their scope. We provide a summary of discrepancies in trait definitions in the appendix (Table \ref{stab:trait_definitions}). Both, differences in differentiation and scope can lead to discrepancies in trait definitions. For example, for the grouping feature feeding mode discrepancies arise because traits are assigned in different ways. The Tachet database defines predators as carvers, engulfers and swallowers. By contrast, in the North American (Twardochleb et al.) database predators are defined as engulfers and carnivorous piercers. In turn, in the Tachet database, piercers are defined as a separate trait encompassing herbivorous and carnivorous piercers. Furthermore, the scope in the freshwaterecology.info database for feeding mode is primarily on the food source of a species (except for filterers), while the other databases focus on the strategies of food acquisition. Therefore, the freshwaterecology.info database defines e.g. predator as "eating from prey", while the other databases use the mouthpart morphology as basis of their definition. The Tachet database captures the food source in an additional grouping feature. Varying levels of differentiation are also present in all other investigated grouping features between the trait databases (Table \ref{tab:trait_databases_coding_differentiation} and Table \ref{stab:trait_definitions}). Locomotion definitions differ also in scope between databases. Freshwaterecology.info and New Zealand databases describe locomotion as the way of movement of an organism, Tachet as substrate relation and locomotion, the North American (Vieira et al.) as how organisms deal with flow, Australia as attachment, and the North American (Twardochleb et al.) database includes not only the way of movement, but also the location of movement. Similarly, regarding reproduction traits, databases differ in their scope. Reproduction is captured in one grouping feature and defined as location of oviposit clutches and mode of reproduction in the freshwaterecology.info and Tachet databases. North America (Vieira et al.) provides information on the oviposition location but not on reproductive behavior. The Australian database report traits for reproductive behavior but also on oviposition site. The New Zealand database distinguishes three grouping features related to reproduction: reproductive technique, oviposition site (e.g. water surface, terrestrial), and egg/egg mass (e.g. free, cemented).

Various codings to describe traits are used throughout the databases (e.g. binary, fuzzy, continuous). The freshwaterecology.info and Australian databases use different codings, while the Tachet and the New Zealand database exclusively use fuzzy coding. Both North American trait databases contain categorical grouping features that can be converted into traits using a binary coding (Table \ref{tab:trait_databases_coding_differentiation}).

\begin{landscape}
  \begin{longtable}{m{2.9cm}|m{1.8cm}|m{2.3cm}|m{1.8cm}|m{3cm}|m{2cm}|m{2cm}|m{1.8cm}}
  \caption{Number of traits per grouping feature and type of coding of the traits for the respective grouping feature per database.}
  \endfirsthead
  \toprule[.1em]
  \label{tab:trait_databases_coding_differentiation}
  Database & Feeding Mode & Voltinism & Locomotion & Respiration & Reproduction & Size & Body Form \\ 
  \toprule[.1em]
  \multirow{2}{*}[-5mm]{ \specialcell{Freshwater- \\ ecology.info}} & 
  10 & 
  6 &
  6 & 
  7 & 
  9 & 
  - & 
  - 
  \\
  \cline{2-8} & 
  \specialcell{10 point \\ assignment \\ system} &
  \specialcell{single category \\ assignment \\ system} &
  \specialcell{10 point \\ assignment \\ system} &
  \multicolumn{2}{c |}{\specialcell{binary}} &
  - & 
  - \\
  \hline
  \hline
  \multirow{2}{*}{Tachet} & 
  7 & 
  3 &
  8 & 
  5 & 
  8 & 
  7 & 
  - 
  \\
  \cline{2-8} &
  \multicolumn{2}{c |}{fuzzy $[0-3]$} &
  fuzzy $[0-5]$ & 
  \multicolumn{3}{c |}{fuzzy $[0-3]$} & 
  -
  \\
  \hline
  \hline
  \multirow{2}{*}{\specialcell{North America \\ (Twardochleb et al.)}} & 
  6 & 
  3 &
  10 & 
  3 & 
  10 & 
  3 & 
  - 
  \\
  \cline{2-8} &
  \multicolumn{6}{c |}{binary} &
  -
  \\
  \hline
  \hline
  \multirow{2}{*}{\specialcell{North America \\ (Vieira et al.)}} & 
  8 & 
  3 &
  9 & 
  8 & 
  10 & 
  3 & 
  5 
  \\
  \cline{2-8} &
  \multicolumn{7}{c }{binary}
  \\
  \hline
  \hline
  \multirow{2}{*}[-4mm]{Australia} & 
  16 \textsuperscript{\textit{a}} & 
  7 &
  9 & 
  10 &
  13 \textsuperscript{\textit{b}} & 
  9 & 
  4
  \\
  \cline{2-8} & 
  \multicolumn{2}{c |}{\specialcell{binary; proportional $[0 - 1]$; \\ fuzzy $[0 - 3]$}} & 
  binary; fuzzy $[0 - 3]$ & 
  \specialcell{binary; proportional \\ scale $[0 - 1]$; \\ fuzzy $[0 - 3]$} & 
  categorical & 
  \specialcell{binary; \\ continuous; \\ fuzzy $[0 - 3]$} & 
  fuzzy codes $[0-3]$
  \\
  \hline
  \hline
  \multirow{2}{*}{New Zealand} & 
  6 & 
  3 & 
  4 & 
  4 & 
  4 & 
  5 & 
  4
  \\
  \cline{2-8} & 
  \multicolumn{7}{c }{fuzzy $[0-3]$}
  \\
  \bottomrule
  \end{longtable}
  \begin{minipage}{\linewidth}\small
      \textit{a} Some of the traits were similar (e.g. trait \textit{Shredder}, \textit{Shredder, Detrivore}, and \textit{Collector, Shredder}).
      \newline
      \textit{b} Not all traits were considered because trait information was partly presented as comments to describe other traits or due to incomplete information.
  \end{minipage}
\end{landscape}

@@

Each trait aggregation method yielded one aggregated affinity for every simulated dataset, in total 15000 comparisons ($100$ replicates $*$ 5 levels of trait variability $* 10$ unique comparisons of trait aggregation methods $* 3$ scenarios). Overall, differences between the results of the aggregation methods for each simulated dataset increased with increasing trait variability. However, in most simulated datasets the different aggregation methods often resulted in similar trait affinities. Only 1.42 \% (213 out of 15.000) of all comparisons showed a difference equal or greater than an absolute trait affinity of 0.1. Most of these differences occurred for the \textit{sim\_extreme} scenario (83.5 \%). The majority of the differences equal or above 0.1 were found between the aggregation methods \textit{direct\_agg\textsubscript{mean}} and \textit{stepwise\_agg\textsubscript{median}}, \textit{direct\_agg\textsubscript{median}} and \textit{stepwise\_agg\textsubscript{median}}, and \textit{stepwise\_agg\textsubscript{median}} and \textit{weighted\_agg} (Figure \ref{fig:sim_indv_runs}). 
% Differences were rather small, largest difference 0.265 for sim_extreme (stepwise_median vs weighted_agg)

@@

Synthesizing trait information from multiple trait databases is crucial for developing the full potential of trait-based approaches, e.g. for studying community trait responses to environmental gradients worldwide. Our attempt at harmonising invertebrate trait information from trait databases of different regions showed that harmonisation was labour-intensive because grouping features are differentiated differently into traits and various codings are used to describe traits in the various databases. In addition, the same traits are sometimes defined differently, requiring expert knowledge to reclassify these traits (e.g. the trait piercer in the Tachet database). Although, others have noted the lack of standardised trait terminology in freshwater ecology (\cite{baird_toward_2011, brink_traits-based_2011}), we are the first to describe invertebrate trait definition discrepancies for some commonly used grouping features. To resolve definition discrepancies and facilitate data synthesis in the future, terminological standards are needed. Harmonised definitions and concepts of traits have been developed in the past for other organism groups, e.g. for plants with the \textit{Thesaurus Of Plant characteristics} (TOP) initiative (\cite{garnier_towards_2017}). The core of this harmonisation initiative is to provide standardised trait definitions and to draw connections to synonyms, related terms, and surrounding concepts by linking to other controlled vocabularies or ontologies in the field. Such initiatives could be a role model for freshwater ecologists to establish unambiguous terminologies for invertebrate traits. The existing freshwater invertebrate trait databases could be linked through standardised terminologies or ontologies, as suggested by \citet{baird_toward_2011} a decade ago. By following the recently proposed \textit{Ecological Trait-data Standard Vocabulary} (ETS), providers of invertebrate trait data could connect their traits to such ontologies (\cite{schneider_towards_2019}). Once standardised trait definitions are established, these will improve trait data sharing, trait data processing when working with multiple trait databases, and ultimately interpretation of the derived results.

@@

When trait values are only available at species-level and observational data are at less precise taxonomic levels, various trait aggregation methods have been used by researchers. We aggregated traits from an Australian and a North American trait dataset to family-level using 5 different aggregation methods and compared the results to traits originally assigned at family-level for these regions. Evaluation of the differences between aggregated and assigned traits is difficult because it remains unclear what the true value of a particular trait for a particular family is. Aggregation of trait information at species or genus-level to a point estimate at family-level suggests a precision that is not necessarily present, especially not for traits with a high variability or if trait information at species-level is missing. Some traits can vary strongly at more precise taxonomic levels than the family-level. For example, \citet{monaghan_improving_2013} found a high intra-family trait dispersion in the Tachet database for traits of the grouping features body size, flow preference, and reproduction. Further, studies focusing on the lability of traits, i.e. how much traits are unconstrained by phylogeny, found that traits of ecological preferences (e.g. thermal preference), body size, resistance forms, and to a lesser extent feeding mode are labile, and thus possibly highly variable (\cite{poff_functional_2006, wilkes_traitbased_2020}). If trait aggregation is necessary, however, our study can give guidance, under which circumstances the choice of aggregation method may be important. The 5 aggregation methods we tested use either the mean or the median and different weightings. Our results indicated that 1) the median aggregation methods compared to mean aggregation methods were often closer to the assigned traits at family-level and 2) the different weightings tested exerted a minor impact on the aggregated trait affinities. We obtained for both datasets more strongly differing cases for mean aggregation methods, irrespective of the weighting. In the Australian dataset, the amount of differing cases was slightly higher for the mean aggregation methods (23 \%) than for median aggregation methods (16 to 18 \% differing cases). In the North American dataset, mean aggregation methods yielded 47 \% cases being different to the assigned traits, a value much higher in comparison to the median aggregation methods (between 15 and 18 \%). When differences occurred between median aggregation methods and assigned traits, these differences were greater compared to the mean aggregation methods, particularly for the North American dataset. There, the mean absolute differences for the median aggregations were twice as high (0.63 - 0.7) compared to the mean aggregations (0.3 - 0.31). These differences can be explained by the binary coding used in the North American trait dataset and by the assigned traits. As a result, traits have affinities of either 1 or 0. Therefore, traits of a particular grouping feature showed in most cases a higher difference in trait affinities to each other than if they were fuzzy coded. Using the median to aggregate binary coded taxa often resulted in a value of 0 or 1, while using the mean yielded values between 0 and 1. Thus, using the median led for the North American dataset either to agreement with the assigned traits or to higher differences compared to the mean aggregation methods.

@@

We explored how using harmonised grouping features and aggregated traits might change the results of trait-environment relationships. We are not aware of other studies that compared trait-environment relationships using harmonised and non harmonised data. The re-analysis of the data on salinisation effects on biological traits from \citet{szocs_effects_2014} yielded only slightly different results compared to the original study, i.e. some of the traits that had responded in the original analysis did not respond in the re-analysis. These non-responsive traits were those closest to the criterion that defined when a trait was associated with either higher or lower salinity (Mahalanobis distance greater than the 97.5 \% quantile of the Chi-square distribution, see method section), namely feeding mode shredder, respiration gills, and life cycle duration traits. Feeding mode shredder was the only trait that was non-responsive when using harmonised (but not aggregated) as well as when using aggregated data. This result was likely due to the harmonisation procedure, when this trait was amalgamated based on three traits (miner, xylophagus, and shredder). Consequently, the trait affinities in the original data had a higher mean and standard deviation than in the aggregated and harmonised data (Table \ref{tab:SI_resp_traits_summary_stats}), suggesting that the signal in the original data was weakened by the harmonisation. %This might be the reason that shredders showed a less strong reaction to the salinity gradient in the re-analysis. -> Kicked out by CM
These findings show that if harmonisation is necessary, harmonised and non-harmonised data, if available, should be compared and possible averaging effects should be considered in further analyses. The fact that the family-level aggregated traits used in the re-analysis showed similar results compared to the original analysis agrees with previous findings that family-level traits can be sensitive enough to detect environmental impacts (\cite{beketov_spear_2009}). 
% Cite another paper that uses family-level trait data
% Although species-level traits are expected to result in more precise results when investigating trait-environment relationships.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% From freshwater biology:
% List all sources in the reference list alphabetically by name. In text citations should follow the author-date method. This means that the author's last name and the year of publication for the source should appear in the text, for example, (Jones, 1998), and a complete reference should appear in the reference list at the end of the paper.

@@

% In press or forthcoming: van Bergen, E., de Jong, P. F., Maassen, B., Krikhaar, E., Plakas, A., & van der Leij, A. (in press). IQ of four-year-olds who go on to develop dyslexia. Journal of Learning Disabilities. DOI: 10.1177/0022219413479673
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

2. We describe discrepancies in the definitions of traits used to create freshwater invertebrate trait databases in Europe, North America, New Zealand, and Australia. Based on our comparisons of these trait databases, we established 4 novel trait datasets by harmonising trait definitions of commonly used traits. Two of these datasets were used to compare aggregated traits obtained by different aggregation methods with traits assigned by experts at family-level. We compared trait aggregation methods that use either the mean or the median and different weightings. Furthermore, we explored the effect of harmonisation and trait aggregation by re-analysing data from a case study.

@@

We established in total 4 trait datasets for the regions Europe, North America, Australia, and New Zealand by harmonising 7 grouping features from 6 trait databases. Trait information for Europe was obtained from the freshwaterecology.info database (\cite{schmidt-kloiber_www.freshwaterecology.info_2015}) and the Tachet database (\cite{ussegliopolatera_biological_2000}). We refer to both databases by their commonly used names (freshwaterecology.info database and Tachet database). The freshwaterecology.info database contains taxa on species-level, while taxa recorded in the Tachet database are on species, genus and family-level. Trait information for North America was obtained from \citet{twardochleb_freshwater_nodate} and complemented by \citet{vieira_database_nodate}. We refer to the former as the CONUS database and the latter as the Vieira database. Data on body form for European and North American taxa was based on expert knowledge (\cite{polatera_personal_information_2020}). For Australia and New Zealand, we used trait databases from \citet{kefford_integrated_2020} and \citet{Philips_and_Smith_NZ_DB_2018} respectively. We refer to both as the Australian and New Zealand database. To increase readability we refer to the datasets we extracted from the trait databases by the name of the continent they originate from.

@@

    \textit{a} The trait "bluff (blocky)" occurred in the Vieira database and was newly classified by expert knowledge into cylindrical and flattened (\cite{polatera_personal_information_2020}). \\  
    \textit{b} Reflects the different size classifications by the Vieiera and the CONUS databases and the other trait databases. \\

@@

\subsubsection*{Simulation of varying taxonomic hierarchies}

To examine the influence of the taxonomic hierarchy and the trait variability on the outcome of the different trait aggregation methods we created three hypothetical examples of different taxonomic hierarchies (scenarios). 
1) A family with an equal number of genera and species (5 genera each with 5 species respectively), denoted as \textit{sim\_base}.
2) A family where one genus has a much larger number of species than the other 4 genera (1 genus with 13 species, 4 genera with 3 species respectively), denoted as \textit{sim\_extreme}. 
3) A family where all genera have a different number of species (8, 2, 7, 3, 5), denoted as \textit{sim\_variation}. In total every family consisted of 25 species. To each scenario a hypothetical grouping feature with 3 traits was assigned (denotes as T1, T2, and T3). The 25 affinities for each trait were simulated by sampling from a truncated normal distribution with a mean value of 0.5 and 5 levels of standard deviations (0.2, 0.4, 0.6, 0.8, and 1) respectively to simulate different levels of trait variability. The truncated normal distribution was bound to 0 and 1. Simulated trait affinities were converted to percentages, similar to the data processing of the trait databases. The sampling was repeated 100 times for each standard deviation. Hence, in total 100 datasets for each level of trait variability were simulated, resulting in 12500 simulated trait affinities for each simulated trait ($25$ species per family $* 5$ levels of trait variability $* 100$ replicates). These 12500 simulated affinities per trait were assigned to each scenario. The above mentioned trait aggregation methods were applied to each simulated dataset. 
% ...resulting in 2500 aggregated affinities per trait and scenario ($12500$ trait affinities per trait divided by $25$, the number of species and multiplied by $5$, the number of aggregation methods). 
The results were compared based on the produced range of aggregated trait affinities between levels of trait variability and scenario. We also compared for each simulated dataset the differences in trait affinities obtained by each aggregation method.

@@

\subsubsection*{Comparison of family-level aggregated traits with family-level assigned traits}

@@

For the North American dataset, we compared aggregated trait affinities with assigned trait affinities on family-level for the grouping features feeding mode, respiration, size, voltinism, and locomotion. The assigned trait affinities at family-level are part of the CONUS database and originate from expert knowledge. Trait information was available for 94 families of which all were present in the North American dataset (total number of cases 1598). The traits were on the categorical scale and were converted to binary traits prior to the comparison with aggregated trait affinities.

@@

Definitions of grouping features and traits varied in their level of detail in the investigated trait databases. The freshwaterecology.info database, the Tachet database and the CONUS database provided more detailed descriptions of their trait information compared to the Vieira database and New Zealand database. The Australian trait database (\cite{kefford_integrated_2020}) is a collection of seven specific trait datasets. Thus, grouping features occur multiple times with varying differentiation into traits. Depending on the dataset trait information is described with more or less detail.

The definition of grouping features varied across databases mainly concerning their differentiation into traits but also in their focus. We provide a summary of discrepancies in trait definitions in the appendix (Table \ref{stab:trait_definitions}). Both, differences in differentiation and scope can lead to discrepancies in trait definitions. For example, for the grouping feature feeding mode discrepancies arise because traits are assigned in different ways. The Tachet database defines predators as carvers, engulfers and swallowers. By contrast, in the CONUS database predators are defined as engulfers and carnivorous piercers. In turn, in the Tachet database, piercers are defined as a separate trait encompassing herbivorous and carnivorous piercers. Furthermore, the focus in the freshwaterecology.info database for feeding mode is primarily on the food source of a species (except for filterers), while the other databases focus on the strategies of food acquisition. Therefore, the freshwaterecology.info database defines e.g. predator as "eating from prey", while the other databases use the mouthpart morphology as basis of their definition. The Tachet database captures the food source in an additional grouping feature. Varying levels of differentiation are also present in all other investigated grouping features between the trait databases (Table \ref{tab:trait_databases_coding_differentiation} and Table \ref{stab:trait_definitions}). Locomotion definitions differ also in focus between databases. Freshwaterecology.info and New Zealand databases describe locomotion as the way of movement of an organism, Tachet as substrate relation and locomotion, the Vieira as how organisms deal with flow, Australia as attachment, and the CONUS database includes not only the way of movement, but also the location of movement. Similarly, regarding reproduction traits, databases differ in their focus. Reproduction is captured in one grouping feature and defined as location of oviposit clutches and mode of reproduction in the freshwaterecology.info and Tachet databases. The Vieira database provides information on the oviposition location but not on reproductive behavior. The Australian database report traits for reproductive behavior but also on oviposition site. The New Zealand database distinguishes three grouping features related to reproduction: reproductive technique, oviposition site (e.g. water surface, terrestrial), and egg/egg mass (e.g. free, cemented).

Various codings to describe traits are used throughout the databases (e.g. binary, fuzzy, continuous). The freshwaterecology.info and Australian databases use different codings, while the Tachet and the New Zealand database exclusively use fuzzy coding. The Vieira and CONUS databases contain categorical grouping features that can be converted into traits using a binary coding (Table \ref{tab:trait_databases_coding_differentiation}).

\begin{landscape}
  \begin{longtable}{m{2.2cm}|m{1.8cm}|m{2.3cm}|m{1.8cm}|m{3cm}|m{2cm}|m{2cm}|m{1.8cm}}
  \caption{Number of traits per grouping feature and type of coding of the traits for the respective grouping feature per database.}
  \endfirsthead
  \toprule[.1em]
  \label{tab:trait_databases_coding_differentiation}
  Database & Feeding Mode & Voltinism & Locomotion & Respiration & Reproduction & Size & Body Form \\ 
  \toprule[.1em]
  \multirow{2}{*}[-5mm]{ \specialcell{Freshwater- \\ ecology.info}} & 
  10 & 
  6 &
  6 & 
  7 & 
  9 & 
  - & 
  - 
  \\
  \cline{2-8} & 
  \specialcell{10 point \\ assignment \\ system} &
  \specialcell{single category \\ assignment \\ system} &
  \specialcell{10 point \\ assignment \\ system} &
  \multicolumn{2}{c |}{\specialcell{binary}} &
  - & 
  - \\
  \hline
  \hline
  \multirow{2}{*}{Tachet} & 
  7 & 
  3 &
  8 & 
  5 & 
  8 & 
  7 & 
  - 
  \\
  \cline{2-8} &
  \multicolumn{2}{c |}{fuzzy $[0-3]$} &
  fuzzy $[0-5]$ & 
  \multicolumn{3}{c |}{fuzzy $[0-3]$} & 
  -
  \\
  \hline
  \hline
  \multirow{2}{*}{\specialcell{CONUS}} & 
  6 & 
  3 &
  10 & 
  3 & 
  10 & 
  3 & 
  - 
  \\
  \cline{2-8} &
  \multicolumn{6}{c |}{binary} &
  -
  \\
  \hline
  \hline
  \multirow{2}{*}{\specialcell{Vieira}} & 
  8 & 
  3 &
  9 & 
  8 & 
  10 & 
  3 & 
  5 
  \\
  \cline{2-8} &
  \multicolumn{7}{c }{binary}
  \\
  \hline
  \hline
  \multirow{2}{*}[-4mm]{Australia} & 
  16 \textsuperscript{\textit{a}} & 
  7 &
  9 & 
  10 &
  13 \textsuperscript{\textit{b}} & 
  9 & 
  4
  \\
  \cline{2-8} & 
  \multicolumn{2}{c |}{\specialcell{binary; proportional $[0 - 1]$; \\ fuzzy $[0 - 3]$}} & 
  binary; fuzzy $[0 - 3]$ & 
  \specialcell{binary; proportional \\ scale $[0 - 1]$; \\ fuzzy $[0 - 3]$} & 
  categorical & 
  \specialcell{binary; \\ continuous; \\ fuzzy $[0 - 3]$} & 
  fuzzy $[0-3]$
  \\
  \hline
  \hline
  \multirow{2}{*}{New Zealand} & 
  6 & 
  3 & 
  4 & 
  4 & 
  4 & 
  5 & 
  4
  \\
  \cline{2-8} & 
  \multicolumn{7}{c }{fuzzy $[0-3]$}
  \\
  \bottomrule
  \end{longtable}
  \begin{minipage}{\linewidth}\small
      \textit{a} Some of the feeding mode traits used in the Australian database were similar (e.g. trait \textit{Shredder}, \textit{Shredder, Detrivore}, and \textit{Collector, Shredder}).
      \newline
      \textit{b} Not all traits were considered because trait information was partly presented as comments to describe other traits or due to incomplete information.
  \end{minipage}
\end{landscape}

@@

Each trait aggregation method yielded one aggregated affinity for every simulated dataset, in total 15000 comparisons ($100$ replicates $*$ 5 levels of trait variability $* 10$ unique comparisons of trait aggregation methods $* 3$ scenarios). In most simulated datasets the different aggregation methods often resulted in similar trait affinities. Only 1.42 \% (213 out of 15.000) of all comparisons showed a difference equal or greater than an absolute trait affinity of 0.1. Most of these differences occurred for the \textit{sim\_extreme} scenario (83.5 \%). The majority of the differences equal or above 0.1 were found between the aggregation methods \textit{direct\_agg\textsubscript{mean}} and \textit{stepwise\_agg\textsubscript{median}}, \textit{direct\_agg\textsubscript{median}} and \textit{stepwise\_agg\textsubscript{median}}, and \textit{stepwise\_agg\textsubscript{median}} and \textit{weighted\_agg} (Figure \ref{fig:sim_indv_runs}). 
% Differences were rather small, largest difference 0.265 for sim_extreme (stepwise_median vs weighted_agg)

@@

Synthesizing trait information from multiple trait databases is crucial for developing the full potential of trait-based approaches, e.g. for studying community trait responses to environmental gradients worldwide. Our attempt at harmonising invertebrate trait information from trait databases of different regions showed that harmonisation was labour-intensive because grouping features are resolved differently into traits and various codings are used to describe traits in the various databases. In addition, the same traits are sometimes defined differently, requiring expert knowledge to reclassify these traits (e.g. the trait piercer in the Tachet database). Although, others have noted the lack of standardised trait terminology in freshwater ecology (\cite{baird_toward_2011, brink_traits-based_2011}), we are the first to describe invertebrate trait definition discrepancies for some commonly used grouping features. To resolve definition discrepancies and facilitate data synthesis in the future, terminological standards are needed. Harmonised definitions and concepts of traits have been developed in the past for other organism groups, e.g. for plants with the \textit{Thesaurus Of Plant characteristics} (TOP) initiative (\cite{garnier_towards_2017}). The core of this harmonisation initiative is to provide standardised trait definitions and to draw connections to synonyms, related terms, and surrounding concepts by linking to other controlled vocabularies or ontologies in the field. Such initiatives could be a role model for freshwater ecologists to establish unambiguous terminologies for invertebrate traits. The existing freshwater invertebrate trait databases could be linked through standardised terminologies or ontologies, as suggested by \citet{baird_toward_2011} a decade ago. By following the recently proposed \textit{Ecological Trait-data Standard Vocabulary} (ETS), providers of invertebrate trait data could connect their traits to such ontologies (\cite{schneider_towards_2019}). Once standardised trait definitions are established, these will improve trait data sharing, trait data processing when working with multiple trait databases, and ultimately interpretation of the derived results.

@@

When trait values are only available at species-level and observational data are at less precise taxonomic levels, various trait aggregation methods have been used by researchers. We aggregated traits from an Australian and a North American trait dataset to family-level using 5 different aggregation methods and compared the results to traits originally assigned at family-level for these regions. Evaluation of the differences between aggregated and assigned traits is difficult because it remains unclear what the true value of a particular trait for a particular family is. Aggregation of trait information at species or genus-level to a point estimate at family-level suggests a precision that is not necessarily present, especially not for traits with a high variability or if trait information at species-level is missing. Some traits can vary strongly at more precise taxonomic levels than the family-level. For example, \citet{monaghan_improving_2013} found a high intra-family trait dispersion in the Tachet database for traits of the grouping features body size, flow preference, and reproduction. Further, studies focusing on the lability of traits, i.e. how much traits are constrained by phylogeny, found that traits of ecological preferences (e.g. thermal preference), body size, resistance forms, and to a lesser extent feeding mode are labile, and thus possibly highly variable (\cite{poff_functional_2006, wilkes_traitbased_2020}). If trait aggregation is necessary, however, our study can give guidance, under which circumstances the choice of aggregation method may be important. The 5 aggregation methods we tested use either the mean or the median and different weightings. Our results indicated that 1) the median aggregation methods compared to mean aggregation methods were often closer to the assigned traits at family-level and 2) the different weightings tested exerted a minor impact on the aggregated trait affinities. We obtained for both datasets more strongly differing cases for mean aggregation methods, irrespective of the weighting. In the Australian dataset, the amount of differing cases was slightly higher for the mean aggregation methods (23 \%) than for median aggregation methods (16 to 18 \% differing cases). In the North American dataset, mean aggregation methods yielded 47 \% cases being different to the assigned traits, a value much higher in comparison to the median aggregation methods (between 15 and 18 \%). When differences occurred between median aggregation methods and assigned traits, these differences were greater compared to the mean aggregation methods, particularly for the North American dataset. There, the mean absolute differences for the median aggregations were twice as high (0.63 - 0.7) compared to the mean aggregations (0.3 - 0.31). These differences can be explained by the binary coding used in the North American trait dataset and by the assigned traits. As a result, traits have affinities of either 1 or 0. Therefore, traits of a particular grouping feature showed in most cases a higher difference in trait affinities to each other than if they were fuzzy coded. Using the median to aggregate binary coded taxa often resulted in a value of 0 or 1, while using the mean yielded values between 0 and 1. Thus, using the median led for the North American dataset either to agreement with the assigned traits or to higher differences compared to the mean aggregation methods.

@@

We explored how using harmonised grouping features and aggregated traits might change the results of trait-environment relationships. We are not aware of other studies that compared trait-environment relationships using harmonised and non harmonised data. The re-analysis of the data on salinisation effects on biological traits from \citet{szocs_effects_2014} yielded only slightly different results compared to the original study, i.e. some of the traits that had responded in the original analysis did not respond in the re-analysis. These non-responsive traits were those closest to the criterion that defined when a trait was associated with either higher or lower salinity, namely feeding mode shredder, respiration gills, and life cycle duration traits. Feeding mode shredder was the only trait that was non-responsive when using harmonised (but not aggregated) as well as when using aggregated data. This result was likely due to the harmonisation procedure, when this trait was amalgamated based on three traits (miner, xylophagus, and shredder). Consequently, the trait affinities in the original data had a higher mean and standard deviation than in the aggregated and harmonised data (Table \ref{tab:SI_resp_traits_summary_stats}), suggesting that the signal in the original data was weakened by the harmonisation. %This might be the reason that shredders showed a less strong reaction to the salinity gradient in the re-analysis. -> Kicked out by CM
These findings show that if harmonisation is necessary, harmonised and non-harmonised data, if available, should be compared and possible averaging effects should be considered in further analyses. The fact that the family-level aggregated traits used in the re-analysis showed similar results compared to the original analysis agrees with previous findings that family-level traits can be sensitive enough to detect environmental impacts (\cite{beketov_spear_2009}). 
% Cite another paper that uses family-level trait data
% Although species-level traits are expected to result in more precise results when investigating trait-environment relationships.
